\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-12-03 main.tex
Minor updates.
Added new file with referee comments

	To proceed with the quantization, we first derive the system's Hamiltonian
	from its corresponding Lagrangian. The Lagrangian for the system is given by
	\begin{align}\label{lagrangian}
		L & = \int \dd^3x \sqrt{-\bar{g}} \left[\frac{\bar{R}}{2\kappa}
			+ \bar{p}(\bar{v}, \bar{s})\right],
	\end{align}
	where, $\bar{g}$ denotes the determinant of the metric, $\bar{R}$ is the background
	Ricci scalar, and $\kappa = \frac{8 \pi G}{c^4}$, where $c$ is the speed of
	light and $G$ is the gravitational constant. We set $c = 1$ for simplicity. The
	term $\bar{p}$ represents the fluid pressure, which is expressed as a function of
	the specific enthalpy $\bar{v}$ and specific entropy $\bar{s}$.

@@

In the above equations, {\color{red} the overdot represents the time derivative and} there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. {\color{red} In our model, the scale factor dynamics resemble
		a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
			\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
			\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
		state $w_q = 1$ that accounts for quantum effects to avoid the primordial
		singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
		physical energy density but rather an effective one that emerges from the Bohmian trajectory. In
		the perturbative analysis, the energy density will always be associated only with
		$\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{b} $,
		Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
		classical dust dominated a flat FLRW universe. With this, we conclude the analysis of our
		background quantum bouncing model and move to its associated perturbations. In the next
		sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and
		$\bar{a}(t_0)/a_b = 10^{35}$ where $t_0$ is the time during the contracting phase where
		$H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover, when solving the equations
		numerically, we use the exact expressions in terms of $w$ without approximation.}

@@

The perturbations are described by the Einstein-Hilbert action expanded up to second order, resulting in the Mukhanov-Sasaki Lagrangian
\begin{align}
	\label{lagms}
	L_{MS} & = \int \dd^3\textbf{x} \frac{1}{ \kappa} \left({\dot{\zeta}}^2 z^2 + c_s^2 z^2 \zeta\Delta \zeta\right)
	,\end{align}
where $\Delta = \bar{D}^2$ is the spatial Laplacian operator and $z$ is given by Eq.~\eqref{zdef}.
To compute the density contrast in Eq.~\eqref{deltarho}, we will need the curvature modes related to the Mukhanov-Sasaki Lagrangian. The extremization of Eq.~\eqref{lagms} in Fourier space yields the equation of motion
\begin{align}
	\label{four}
	\Ddot{\zeta}_{k}+ 2\frac{ \dot{z} }{ z } \dot{\zeta}_k + \frac{k^2}{\bar{a}^2} \zeta_k & =0
	,\end{align}
where $k$ is the comoving wave number of the modes. Since obtaining analytical solutions to \eqref{four} is non-trivial, we will employ a numerical code to compute the modes, which requires the use of dimensionless variables. In dimensionless units, denoted with a subscript $A$, we redefine our variables as\footnote{A derivative of a dimensionless variable will also be dimensionless. However, we use the same dot notation.}
\begin{align}
	\zeta_{k_A, A}   & \equiv \frac{\zeta_{k}}{\sqrt{ \kappa \hbar R_H}},           \\
	\Pi_{\zeta_k, A} & \equiv \frac{\Pi_{\zeta_k} \sqrt{R_H}}{\sqrt{\kappa \hbar}}, \\
	\label{kadm}
	k_A              & \equiv k R_H,                                                \\
	t_A              & \equiv \frac{t}{R_H}
	,\end{align}
where $l_p$ is the Planck length and $R_H = \frac{1}{\bar{H}_0}$. The equation of motion becomes
\begin{align}
	\label{eomosc}
	\dot{\Pi}_{k_A, A}+  \frac{2 c^2_s z^2 k_A^2}{\bar{a}^2  } \zeta_{k_A, A} & =0 \nonumber~~ or \\
	\dot{\Pi}_{k_A, A}+  2 \omega_{k_a}^2 z^2  \zeta_{k_A, A}                 & =0
	.\end{align}
Here, $\omega_A^2 \equiv \frac{c_s^2 k_a^2}{\bar{a}^2}$ is the dimensionless frequency and the dimensionless conjugated momenta are given by
\begin{align}
	\Pi_{k_A, A} & = 2 z^2 \dot{\zeta}_{k_A, A}
	.\end{align}
Eq.~\eqref{eomosc} can be interpreted as a harmonic oscillator with a time-dependent frequency $\omega_{k_a}(t_a)$ and mass $m(t_a) = 2z^2$. From here on, we are going to use the dimensionless variables with the same notation as described above.

@@

\textcolor{red}{In this work, we have investigated the formation of primordial black
	holes within a Friedmann background, incorporating scalar perturbations. An alternative
	approach would be to explore PBH formation in a homogeneous yet anisotropic background.
	The presence of anisotropy could give rise to PBHs with significant angular momentum,
	contrasting with those formed in standard inflationary scenarios. A study of this
	possibility is left for future work.}. The next phase of this research would be to apply
the same methodology to a universe populated with radiation and dark matter, which more
closely resembles the physical universe and the interaction between both fluids may
affect PBH production. If such a model proves reasonable, it would also be interesting
to improve the computation of the LTB model in App.~\ref{appc} for larger values of the
equation of state. Despite the Jeans scale being a good approximation, the introduction
of a non-pressureless fluid would also require the study of the pressure impact in our
characteristic scales. Furthermore, in a radiation-dominated quantum bouncing model, one
expects a greater blue tilt on the spectral index, which may affect the density contrast
modes for smaller scales. Said effect could lead to a larger formation of PBHs. This
analysis is left for future work.

>>>>

	To proceed with the quantization, we first derive the system's Hamiltonian
	from its corresponding Lagrangian. The Lagrangian of the system is given by
	\begin{align}\label{lagrangian}
		L & = \int \dd^3x \sqrt{-\bar{g}} \left[\frac{\bar{R}}{2\kappa}
			+ \bar{p}(\bar{v}, \bar{s})\right],
	\end{align}
	where, $\bar{g}$ denotes the determinant of the metric, $\bar{R}$ is the background
	Ricci scalar, and $\kappa = \frac{8 \pi G}{c^4}$, where $c$ is the speed of
	light and $G$ is the gravitational constant. We set $c = 1$ for simplicity. The
	term $\bar{p}$ represents the fluid pressure, which is expressed as a function of
	the specific enthalpy $\bar{v}$ and specific entropy $\bar{s}$.

@@

In the above equations, {\color{red} the overdot represents the time derivative and} there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. {\color{red} In our model, the scale factor dynamics resemble
		a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
			\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
			\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
		state $w_q = 1$ that accounts for quantum effects to avoid the primordial
		singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
		physical energy density but rather an effective one that emerges from the Bohmian trajectory. In
		the perturbative analysis, the energy density will always be associated only with
		$\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{B} $,
		Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
		classical dust dominated a flat FLRW universe. With this, we conclude the analysis of our
		background quantum bouncing model and move to its associated perturbations. In the next
		sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and
		$\bar{a}(t_0)/a_B = 10^{35}$ where $t_0$ is the time during the contracting phase where
		$H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover, when solving the equations
		numerically, we use the exact expressions in terms of $w$ without approximation.}

@@

The perturbations are described by the Einstein-Hilbert action expanded up to second order, resulting in the Mukhanov-Sasaki Lagrangian
\begin{align}
	\label{lagms}
	L_{MS} & = \int \dd^3\textbf{x} \frac{1}{ \kappa} \left({\dot{\zeta}}^2 z^2 + c_s^2 z^2 \zeta\Delta \zeta\right)
	,\end{align}
where $\Delta = \bar{D}^2$ is the spatial Laplacian operator and $z$ is given by Eq.~\eqref{zdef}.
To compute the density contrast in Eq.~\eqref{deltarho}, we will need the curvature modes related to the Mukhanov-Sasaki Lagrangian. The extremization of Eq.~\eqref{lagms} in Fourier space yields the equation of motion
\begin{align}
	\label{four}
	\Ddot{\zeta}_{k}+ 2\frac{ \dot{z} }{ z } \dot{\zeta}_k + \frac{k^2}{\bar{a}^2} \zeta_k & =0
	,\end{align}
where $k$ is the comoving wave number of the modes. Since obtaining analytical solutions to \eqref{four} is non-trivial, we will employ a numerical code to compute the modes, which requires the use of dimensionless variables. In dimensionless units, denoted with a subscript $A$, we redefine our variables as\footnote{A derivative of a dimensionless variable will also be dimensionless. However, we use the same dot notation.}
\begin{align}
	\zeta_{k_A, A}   & \equiv \frac{\zeta_{k}}{\sqrt{ \kappa \hbar R_H}},           \\
	\Pi_{\zeta_k, A} & \equiv \frac{\Pi_{\zeta_k} \sqrt{R_H}}{\sqrt{\kappa \hbar}}, \\
	\label{kadm}
	k_A              & \equiv k R_H,                                                \\
	t_A              & \equiv \frac{t}{R_H}
	,\end{align}
where $R_H = \frac{1}{\bar{H}_0}$. The equation of motion becomes
\begin{align}
	\label{eomosc}
	\dot{\Pi}_{k_A, A}+  \frac{2 c^2_s z^2 k_A^2}{\bar{a}^2  } \zeta_{k_A, A} & =0 \nonumber~~ or \\
	\dot{\Pi}_{k_A, A}+  2 \omega_{k_a}^2 z^2  \zeta_{k_A, A}                 & =0
	.\end{align}
Here, $\omega_A^2 \equiv \frac{c_s^2 k_a^2}{\bar{a}^2}$ is the dimensionless frequency and the dimensionless conjugated momenta are given by
\begin{align}
	\Pi_{k_A, A} & = 2 z^2 \dot{\zeta}_{k_A, A}
	.\end{align}
Eq.~\eqref{eomosc} can be interpreted as a harmonic oscillator with a time-dependent frequency $\omega_{k_a}(t_a)$ and mass $m(t_a) = 2z^2$. From here on, we are going to use the dimensionless variables with the same notation as described above.

@@

\textcolor{red}{In this work, we have investigated the formation of primordial black
	holes within a Friedmann background, incorporating scalar perturbations. An alternative
	approach would be to explore PBH formation in a homogeneous yet anisotropic background.
	The presence of anisotropy could give rise to PBHs with significant angular momentum,
	contrasting with those formed in standard inflationary scenarios. A study of this
	possibility is left for future work}. The next phase of this research would be to apply
the same methodology to a universe populated with radiation and dark matter, which more
closely resembles the physical universe and the interaction between both fluids may
affect PBH production. If such a model proves reasonable, it would also be interesting
to improve the computation of the LTB model in App.~\ref{appc} for larger values of the
equation of state. Despite the Jeans scale being a good approximation, the introduction
of a non-pressureless fluid would also require the study of the pressure impact in our
characteristic scales. Furthermore, in a radiation-dominated quantum bouncing model, one
expects a greater blue tilt on the spectral index, which may affect the density contrast
modes for smaller scales. Said effect could lead to a larger formation of PBHs. This
analysis is left for future work.
